Hilbert space methods to approximate Gaussian processes using Stan

### 2019-03-04 Paper/manuscript/manuscript_05.tex
spell check

Yet another approach to approximate GPs -- on which we will focus in the present paper -- was proposed by \cite{solin2018hilbert} who suggested to use reduced-rank approximations of the covariance kernel. This method decomposes the kernel into basis functions which represent the GP as a linear model and thus makes its estimation considerably faster. In principal, we are free in the choice of basis functions, but the Laplace eigenfunctions were found to be particularily appealing \citep{solin2018hilbert}. Not only can they be computed analytically, but they are also independent of the particular choice of the covariance kernel including the hyperparameters. The latter comes with several advantages when performing inference via algorithms that require differentiation of the posterior density (Paul: cite something here). In a fully Bayesian framework based on sampling methods, the proposed approximate GP model has a computational complexity of just $O(nm+m)$, with $m<<n$ in every sampling step. The posterior distribution of the reduced-rank GP approximation is of dimension $m$ and thus considerably smaller than in the case of exact GPs. This not only lowers the memory requirements and computational complexity but also reduces the correlations between latent values and the corresponding hyperparameters of the GP, which can help to further improve sampling efficiency.

@@

The remainder of the paper is structered as follows. In Section 2, we introduce GPs and their reduced rank approximations proposed by \cite{solin2018hilbert}. In Section 3, we analyze the accuracy of these approximations under several conditions using analytical and numerical methods. Several case studies in which we fit
exact and approximate GPs to real and simulated data are provided in Section 4. We end with a discussion in Section 5.

@@

\noindent where $\boldsymbol{\mu} = \left\lbrace \mu(\mathbf{x}_n) \right\rbrace_{n=1}^N$ is the multivariate mean and $K$ the covariance matrix, where the element $K_{i,j}$ are formed by evalating the covariance function $k(\mathbf{x}_i,\mathbf{x}_j)$ at input values $\mathbf{x}_i$ and $\mathbf{x}_j$. The joint distribution of $\mathbf{f}$ and $f^*$ is also a multivariate Gaussian as,
%
\begin{eqnarray*}
p(\mathbf{f},\mathbf{f}')=\mathcal{N} \left( \left[ \begin{array}{cc}
\mathbf{f} \\ 
\mathbf{f}'
\end{array} \right] | 0,\left[ \begin{array}{cc}
K_{\mathbf{f},\mathbf{f}} & K_{\mathbf{f},f^*} \\ 
K_{f^*,\mathbf{f}} & K_{f^*,f^*}
\end{array} \right] \right),
\end{eqnarray*}

@@

In most modelling problems, however, the observational model is not Gaussian (e.g. binomial or multinomial classification, Poisson  regression, and many other modelling problems) and the posterior predictive distribution $p(f^*| \mathbf{y})$ is analytically untractable and can not be obtained in a closed-form expression. Approximate inference have to be used to tackle the intractable inference, tipically based on sampling methods such as MCMC \citep{brooks_2011} or analytical Gaussian approximations such as Laplace approximation (\citep{williams1998bayesian,rasmussen2006gaussian}, expectation propagation \citep{minka2001expectation}, or variational methods \citep{gibbs2000variational,csato2000efficient}. Furthermore, Gaussian processes in a direct implementation scales cubically in the number of sample points, $N$, which effectively limits the modelling inference for medium and large datasets. In the introduction section in this paper several methods tipically used for dealing with the computational complexity has been commented. \\

@@

The covariance function is the crucial ingredient in a Gaussian process as it encodes our prior assumptions about the function, and defines a correlation structure which characterize the correlations between function values at different inputs. A stationary covariance function is a function of $\mathbf{x}-\mathbf{x}'$, such that it can be written $k(\mathbf{x},\mathbf{x}') = k(|\mathbf{x}-\mathbf{x}'|)$, which means that the covariance is invariant to translations. Isotropic covariance functions are those that are function of the distance between observations, $k(\mathbf{x},\mathbf{x}') = k(||\mathbf{x}-\mathbf{x}'||) = k(\mathbf{r})$, which means that the covariacne is both translation and rotation invariant. The Mat\'ern family with half-integers is the class of isotropic covariance functions probably most commonly used, which, in the unidimensional case (i.e. $\mathbf{x}=x\in \mathbb{R}$, so $\mathbf{r}=r\in \mathbb{R}$), are
%
\begin{eqnarray}
k_{\nu=\infty}(r)&=&\sigma^2 \text{exp}(-\frac{1}{2} r^2/\ell^2), \nonumber \\
k_{\nu=\frac{1}{2}}(r)&=&\sigma^2 exp(-r/\ell), \nonumber \\
k_{\nu=\frac{3}{2}}(r)&=&\sigma^2(1+\sqrt{3}r/\ell \text{exp}(-\sqrt{3}r/\ell), \nonumber \\
k_{\nu=\frac{5}{2}}(r)&=&\sigma^2(1+\sqrt{5}r/\ell+\frac{5}{3}r^2/\ell^2) \text{exp}(-\sqrt{5}r/\ell), \nonumber 
\end{eqnarray}

\noindent where $\nu$ is the order the kernel, and the $\ell$ and $\sigma$ are the lengt-scale and magnitud, respectively, of the kernel. The particular case where $\nu=\infty$ is commonly known as squared exponential (exponentiated quadratic) covariance function, and that with $\nu=1/2$ as exponential covariance function. These covariance functions can be easily generalized to the multidimensional case, with multidimensional inputs $\mathbf{x}\in \mathbb{R}^D$ and multidimensional length-scale $\boldsymbol{\ell}\in \mathbb{R}^D$ (Gabi: We should derive the multidimensional form of the previous kernels, but I have now doubts how is the multidimensional form for the Matern kernels with $\nu$=1/2, $\nu$=3/2 and $\nu$=5/2). Using a multidimensional length-scale basically turns the isotropic covariance function into non-isotropic.

@@

We begin by focusing on the case of a unidimensional input space (i.e., on GPs with just a single covariate) such that $\Omega \in \{-L,L\} \subset {\rm I\!R}$, where $L$ is some positive real value to which we also refer as boundary condition. As $\Omega$ describes the interval in which the approximations are valid, $L$ plays a critical role in the accuarcy of HSGPs. We will come back to this issue in Section 3.

@@

\vspace{2mm}
$\bullet$ In a fully Bayesian inference framework using sampling methods, the proposed approximate GP model has a computational complexity of $O(nm+m)$ in every step of the HMC method. In addition, the computation of the automatic differentation to compute the gradients in this linear model scales $O(n)$?, an operation that must be computed in every step of the HMC method.

@@

\vspace{2mm}
$\bullet$ In conventional sparse GP approximations, although the rank of the GP is reduced considerably to the number of inducing points, this still needs to do the autodiff and covariane matrix inversion.

\vspace{2mm}
$\bullet$ The Splines models are also a sort of basis functions expansion model, then the computational demands are similar to that in this approach. However in Splines models the lengthscale hyperparameter tend to be fixed and then the fit is covered by the magnitud parameter. In that sense, Splines models tend to loose the useful interpretation of the lengthscale parameter.

\vspace{2mm}
$\bullet$ In addition, the computation of the automatic differentation to compute gradients in this linear model scales $O(n)$, which is an operation that must be computed in every step of the HMC method.

\vspace{2mm}
$\bullet$ In a regular GP model the automatic differentation to compute the gradients of the covariance function scales $O(n^2)$, the dimension of the covariance matrix, and the full inversion of the covariance matrix scales $O(n^3)$. This operation has to be computed at every step of the HMC.

\vspace{2mm}
$\bullet$ In a sparse GP approach based on inducing points, although the rank of the GP is reduced considerably to the number of inducing points, this still needs to do the autodiff and covariane matrix inversion.

@@

For a naive illustration of the effects of these two factors, number of basis functions $m$ and boundary $L$, on the performance of the approximation, a set of noisy observations are drawn from a GP model with lengthscale $\ell=0.3$ and marginal variance $\alpha=1$, and input vales within the zero-mean input domain with half-range $S=1$. HSGP models with different number of basis functions and boundaries are fitted to this data. The lengthscale and marginal variance parameters are considered fixed to the true values of the data-generative GP model. Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} ilustrate the individual effects of the number of basis functions and the boundary, respectively, on the approximation of the covariance function. On the one hand, for a fixed $L$, let's suppose it is set to a right choice, Figure \ref{fig1_Post_J} shows the effect of changing the number of basis functions in approaching the posterior mean and the covariance function. It can be seen how the number of basis functions affects the non-linear effects of the posteriors. Fewer basis functions implies smoother covariance functions and consequently smoother posteriors. The more "wigglyness" of the function, the more basis functions are needed. On the other hand, and similarly, for a fixed number of basis functions $m$, let's suppose it is set to a right choice, Figure \ref{fig2_Post_L} shows the effect of changing the boundary $L$ in approaching the posterior mean and the covariance function. It can be seen how the box size mainly affects the approximation near the boundaries of the function and affects the covariances at long distances. \\

@@

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.4]{fig5_MSE_vs_J.pdf}}
\subfigure{\includegraphics[scale=0.4]{fig5_MSE_vs_c.pdf}}
\caption{Standarized root mean square error (SRMSE) of the proposed HSGP models computed against the regular GP model. (left) SRMSE vs the number of basis functions $m$ and for different values of the boundary factor $c$. (right) SRMSE vs the boundary factor $c$ and for different values of the number of basis functions $m$. }
  \label{fig5_MSE_vs_J}
\end{figure}

@@

$\bullet$ This plot serves as a diagnosis tool in the sense that if a estimated lengthscale is lower than that minimun that the figure bring us, given certain number of basis functions and certain boundary factor, means that the number of basis functions should be increased or the boundary factor decreased. On the other hand, if the lengthscale is bigger than that generated by the figure means that the approximation should be close enough.

@@

\begin{figure}[H]
\centering
\subfigure{\includegraphics[scale=0.4]{fig3_lscale_vs_J_vs_c_part1.pdf}}
\subfigure{\includegraphics[scale=0.4]{fig3_lscale_vs_J_vs_c_part2.pdf}}
\caption{Relation between the minumum number of basis functions $m$, the lengthscale normalized by the half-range of the data ($\frac{\ell}{S}$), and the boundary factor $c$ ($c = \frac{L}{S}$).}
  \label{fig3_lscale_vs_J_vs_c_part1}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{fig8_Tlscale_vs_Elscale.pdf}
\caption{True lengthscales versus estimated lscales of the different realizations of datasets using different number of basis functions. }
  \label{fig8_Tlscale_vs_Elscale}
\end{figure}

@@

In order to do model comparison, in addition to the regular GP model and HSGP model, an splines-based model is also fitted using the Thin Plate Regression Splinesn approach in \cite{wood2003thin} and implemented in the R-package \textit{mgcv}. A Bayesian approach is used to fit this splines model using the R-package \textit{brms}.

@@

\begin{figure}[H]
\centering
\subfigure{\includegraphics[scale=0.350]{fig11_MSE_exI_inter.pdf}}
\subfigure{\includegraphics[scale=0.350]{fig11_MSE_exI_extra.pdf}}
\caption{Standarized mean square error (SMSE) of the different methods againts the true function. (left) SMSE for interpolation. (right) SMSE for extrapolation}
  \label{fig11_MSE_exI_inter}
\end{figure}

@@

\noindent with a squared exponetial covariance function $k$ (2).

@@

\begin{figure}[H]
\centering
\subfigure{\includegraphics[scale=0.40]{fig13_MSE_train_BF_gaydata.pdf}}
\subfigure{\includegraphics[scale=0.40]{fig13_MSE_pred_BF_gaydata.pdf}}
\caption{Mean square error (MSE) of the different methods againts the true function. (left) MSE for training data. (right) MSE for test data.}
  \label{fig13_MSE_train_BF_gaydata}
\end{figure}

@@

\begin{figure}[H]
\centering
\subfigure{\includegraphics[scale=0.40]{fig14_MSE_train_gaydata.pdf}}
\subfigure{\includegraphics[scale=0.40]{fig14_MSE_pred_gaydata.pdf}}
\caption{Mean square error (MSE) of the different methods againts the true function. (left) MSE for training data. (right) MSE for test data.}
  \label{fig14_MSE_train_gaydata}
\end{figure}

@@

This scheme was originally introduced to the GP context by \cite{williams2001using}. As discussed by \cite{quinonero2005unifying}, the Nystr\"om method by \cite{williams2001using} does not correspond to a well-formed probabilistic model. However, several methods modifying the inducing point approach are widely used. The Subset of Regressors (SoR) \citep{smola2001sparse} method uses the Nystr\"om approximation scheme and a finite linear-in-the-parameters model for approximating the whole (training and test) covariance function, whereas the sparse Nystr\"om method \citep{williams2001using} only replaces the training data covariance matrix. The SoR method is based on a degenarete prior which produces unresoanble predictive uncertainties, which is a general problem of linear models (for more details see \cite{rasmussen2006gaussian}).

@@

\vspace{2mm}
$\bullet$ In conventional sparse GP approximations, although the rank of the GP is reduced considerably to the number of inducing points, this still needs to do the autodiff and covariane matrix inversion.

\vspace{2mm}
$\bullet$ The Splines models are also a sort of basis functions expansion model, then the computational demands are similar to that in this approach. However in Splines models the lengthscale hyperparameter tend to be fixed and then the fit is covered by the magnitud parameter. In that sense, Splines models tend to loose the useful interpretation of the lengthscale parameter.

@@

\vspace{2mm}
$\bullet$ We investigate the relations going on among these factors, the number of basis functions, the box size, and the lengthsccale of the functions.

\vspace{2mm}
$\bullet$ We make recomendations for the values of these factors based on the recognized relations among them. We provide useful graphs of these relations that will help the users to improve performance and save time of computation.

@@

Assuming that the espectral density function $S(\cdot)$ is regular enough, then it can be represented as a polynomial expansion:
%
\begin{equation}
S(\mathbf{w})=a_0+a_1\mathbf{w}^2+a_2(\mathbf{w}^2)^2+a_1(\mathbf{w}^2)^3+\cdots
\end{equation}

>>>>

Yet another approach to approximate GPs -- on which we will focus in the present paper -- was proposed by \cite{solin2018hilbert} who suggested to use reduced-rank approximations of the covariance kernel. This method decomposes the kernel into basis functions which represent the GP as a linear model and thus makes its estimation considerably faster. In principal, we are free in the choice of basis functions, but the Laplace eigenfunctions were found to be particularly appealing \citep{solin2018hilbert}. Not only can they be computed analytically, but they are also independent of the particular choice of the covariance kernel including the hyperparameters. The latter comes with several advantages when performing inference via algorithms that require differentiation of the posterior density (Paul: cite something here). In a fully Bayesian framework based on sampling methods, the proposed approximate GP model has a computational complexity of just $O(nm+m)$, with $m<<n$ in every sampling step. The posterior distribution of the reduced-rank GP approximation is of dimension $m$ and thus considerably smaller than in the case of exact GPs. This not only lowers the memory requirements and computational complexity but also reduces the correlations between latent values and the corresponding hyperparameters of the GP, which can help to further improve sampling efficiency.

@@

The remainder of the paper is structured as follows. In Section 2, we introduce GPs and their reduced rank approximations proposed by \cite{solin2018hilbert}. In Section 3, we analyze the accuracy of these approximations under several conditions using analytical and numerical methods. Several case studies in which we fit
exact and approximate GPs to real and simulated data are provided in Section 4. We end with a discussion in Section 5.

@@

\noindent where $\boldsymbol{\mu} = \left\lbrace \mu(\mathbf{x}_n) \right\rbrace_{n=1}^N$ is the multivariate mean and $K$ the covariance matrix, where the element $K_{i,j}$ are formed by evaluating the covariance function $k(\mathbf{x}_i,\mathbf{x}_j)$ at input values $\mathbf{x}_i$ and $\mathbf{x}_j$. The joint distribution of $\mathbf{f}$ and $f^*$ is also a multivariate Gaussian as,
%
\begin{eqnarray*}
p(\mathbf{f},\mathbf{f}')=\mathcal{N} \left( \left[ \begin{array}{cc}
\mathbf{f} \\ 
\mathbf{f}'
\end{array} \right] | 0,\left[ \begin{array}{cc}
K_{\mathbf{f},\mathbf{f}} & K_{\mathbf{f},f^*} \\ 
K_{f^*,\mathbf{f}} & K_{f^*,f^*}
\end{array} \right] \right),
\end{eqnarray*}

@@

In most modelling problems, however, the observational model is not Gaussian (e.g. binomial or multinomial classification, Poisson  regression, and many other modelling problems) and the posterior predictive distribution $p(f^*| \mathbf{y})$ is analytically untractable and can not be obtained in a closed-form expression. Approximate inference have to be used to tackle the intractable inference, typically based on sampling methods such as MCMC \citep{brooks_2011} or analytical Gaussian approximations such as Laplace approximation (\citep{williams1998bayesian,rasmussen2006gaussian}, expectation propagation \citep{minka2001expectation}, or variational methods \citep{gibbs2000variational,csato2000efficient}. Furthermore, Gaussian processes in a direct implementation scales cubically in the number of sample points, $N$, which effectively limits the modelling inference for medium and large datasets. In the introduction section in this paper several methods typically used for dealing with the computational complexity has been commented. \\

@@

The covariance function is the crucial ingredient in a Gaussian process as it encodes our prior assumptions about the function, and defines a correlation structure which characterize the correlations between function values at different inputs. A stationary covariance function is a function of $\mathbf{x}-\mathbf{x}'$, such that it can be written $k(\mathbf{x},\mathbf{x}') = k(|\mathbf{x}-\mathbf{x}'|)$, which means that the covariance is invariant to translations. Isotropic covariance functions are those that are function of the distance between observations, $k(\mathbf{x},\mathbf{x}') = k(||\mathbf{x}-\mathbf{x}'||) = k(\mathbf{r})$, which means that the covariance is both translation and rotation invariant. The Mat\'ern family with half-integers is the class of isotropic covariance functions probably most commonly used, which, in the unidimensional case (i.e. $\mathbf{x}=x\in \mathbb{R}$, so $\mathbf{r}=r\in \mathbb{R}$), are
%
\begin{eqnarray}
k_{\nu=\infty}(r)&=&\sigma^2 \text{exp}(-\frac{1}{2} r^2/\ell^2), \nonumber \\
k_{\nu=\frac{1}{2}}(r)&=&\sigma^2 exp(-r/\ell), \nonumber \\
k_{\nu=\frac{3}{2}}(r)&=&\sigma^2(1+\sqrt{3}r/\ell \text{exp}(-\sqrt{3}r/\ell), \nonumber \\
k_{\nu=\frac{5}{2}}(r)&=&\sigma^2(1+\sqrt{5}r/\ell+\frac{5}{3}r^2/\ell^2) \text{exp}(-\sqrt{5}r/\ell), \nonumber 
\end{eqnarray}

\noindent where $\nu$ is the order the kernel, and the $\ell$ and $\sigma$ are the length-scale and magnitude, respectively, of the kernel. The particular case where $\nu=\infty$ is commonly known as squared exponential (exponentiated quadratic) covariance function, and that with $\nu=1/2$ as exponential covariance function. These covariance functions can be easily generalized to the multidimensional case, with multidimensional inputs $\mathbf{x}\in \mathbb{R}^D$ and multidimensional length-scale $\boldsymbol{\ell}\in \mathbb{R}^D$ (Gabi: We should derive the multidimensional form of the previous kernels, but I have now doubts how is the multidimensional form for the Matern kernels with $\nu$=1/2, $\nu$=3/2 and $\nu$=5/2). Using a multidimensional length-scale basically turns the isotropic covariance function into non-isotropic.

@@

We begin by focusing on the case of a unidimensional input space (i.e., on GPs with just a single covariate) such that $\Omega \in \{-L,L\} \subset {\rm I\!R}$, where $L$ is some positive real value to which we also refer as boundary condition. As $\Omega$ describes the interval in which the approximations are valid, $L$ plays a critical role in the accuracy of HSGPs. We will come back to this issue in Section 3.

@@

\vspace{2mm}
$\bullet$ In a fully Bayesian inference framework using sampling methods, the proposed approximate GP model has a computational complexity of $O(nm+m)$ in every step of the HMC method. In addition, the computation of the automatic differentiation to compute the gradients in this linear model scales $O(n)$?, an operation that must be computed in every step of the HMC method.

@@

\vspace{2mm}
$\bullet$ In conventional sparse GP approximations, although the rank of the GP is reduced considerably to the number of inducing points, this still needs to do the autodiff and covariance matrix inversion.

\vspace{2mm}
$\bullet$ The Splines models are also a sort of basis functions expansion model, then the computational demands are similar to that in this approach. However in Splines models the lengthscale hyperparameter tend to be fixed and then the fit is covered by the magnitude parameter. In that sense, Splines models tend to loose the useful interpretation of the lengthscale parameter.

\vspace{2mm}
$\bullet$ In addition, the computation of the automatic differentiation to compute gradients in this linear model scales $O(n)$, which is an operation that must be computed in every step of the HMC method.

\vspace{2mm}
$\bullet$ In a regular GP model the automatic differentiation to compute the gradients of the covariance function scales $O(n^2)$, the dimension of the covariance matrix, and the full inversion of the covariance matrix scales $O(n^3)$. This operation has to be computed at every step of the HMC.

\vspace{2mm}
$\bullet$ In a sparse GP approach based on inducing points, although the rank of the GP is reduced considerably to the number of inducing points, this still needs to do the autodiff and covariance matrix inversion.

@@

For a naive illustration of the effects of these two factors, number of basis functions $m$ and boundary $L$, on the performance of the approximation, a set of noisy observations are drawn from a GP model with lengthscale $\ell=0.3$ and marginal variance $\alpha=1$, and input vales within the zero-mean input domain with half-range $S=1$. HSGP models with different number of basis functions and boundaries are fitted to this data. The lengthscale and marginal variance parameters are considered fixed to the true values of the data-generative GP model. Figures \ref{fig1_Post_J} and \ref{fig2_Post_L} illustrate the individual effects of the number of basis functions and the boundary, respectively, on the approximation of the covariance function. On the one hand, for a fixed $L$, let's suppose it is set to a right choice, Figure \ref{fig1_Post_J} shows the effect of changing the number of basis functions in approaching the posterior mean and the covariance function. It can be seen how the number of basis functions affects the non-linear effects of the posteriors. Fewer basis functions implies smoother covariance functions and consequently smoother posteriors. The more "wigglyness" of the function, the more basis functions are needed. On the other hand, and similarly, for a fixed number of basis functions $m$, let's suppose it is set to a right choice, Figure \ref{fig2_Post_L} shows the effect of changing the boundary $L$ in approaching the posterior mean and the covariance function. It can be seen how the box size mainly affects the approximation near the boundaries of the function and affects the covariances at long distances. \\

@@

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.4]{fig5_MSE_vs_J.pdf}}
\subfigure{\includegraphics[scale=0.4]{fig5_MSE_vs_c.pdf}}
\caption{Standardized root mean square error (SRMSE) of the proposed HSGP models computed against the regular GP model. (left) SRMSE vs the number of basis functions $m$ and for different values of the boundary factor $c$. (right) SRMSE vs the boundary factor $c$ and for different values of the number of basis functions $m$. }
  \label{fig5_MSE_vs_J}
\end{figure}

@@

$\bullet$ This plot serves as a diagnosis tool in the sense that if a estimated lengthscale is lower than that minimum that the figure bring us, given certain number of basis functions and certain boundary factor, means that the number of basis functions should be increased or the boundary factor decreased. On the other hand, if the lengthscale is bigger than that generated by the figure means that the approximation should be close enough.

@@

\begin{figure}[H]
\centering
\subfigure{\includegraphics[scale=0.4]{fig3_lscale_vs_J_vs_c_part1.pdf}}
\subfigure{\includegraphics[scale=0.4]{fig3_lscale_vs_J_vs_c_part2.pdf}}
\caption{Relation between the minimum number of basis functions $m$, the lengthscale normalized by the half-range of the data ($\frac{\ell}{S}$), and the boundary factor $c$ ($c = \frac{L}{S}$).}
  \label{fig3_lscale_vs_J_vs_c_part1}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{fig8_Tlscale_vs_Elscale.pdf}
\caption{True lengthscales versus estimated lengthscales of the different realizations of datasets using different number of basis functions. }
  \label{fig8_Tlscale_vs_Elscale}
\end{figure}

@@

@@

\begin{figure}[H]
\centering
\subfigure{\includegraphics[scale=0.350]{fig11_MSE_exI_inter.pdf}}
\subfigure{\includegraphics[scale=0.350]{fig11_MSE_exI_extra.pdf}}
\caption{Standardized mean square error (SMSE) of the different methods against the true function. (left) SMSE for interpolation. (right) SMSE for extrapolation}
  \label{fig11_MSE_exI_inter}
\end{figure}

@@

\noindent with a squared exponential covariance function $k$ (2).

@@

\begin{figure}[H]
\centering
\subfigure{\includegraphics[scale=0.40]{fig13_MSE_train_BF_gaydata.pdf}}
\subfigure{\includegraphics[scale=0.40]{fig13_MSE_pred_BF_gaydata.pdf}}
\caption{Mean square error (MSE) of the different methods against the true function. (left) MSE for training data. (right) MSE for test data.}
  \label{fig13_MSE_train_BF_gaydata}
\end{figure}

@@

\begin{figure}[H]
\centering
\subfigure{\includegraphics[scale=0.40]{fig14_MSE_train_gaydata.pdf}}
\subfigure{\includegraphics[scale=0.40]{fig14_MSE_pred_gaydata.pdf}}
\caption{Mean square error (MSE) of the different methods against the true function. (left) MSE for training data. (right) MSE for test data.}
  \label{fig14_MSE_train_gaydata}
\end{figure}

@@

This scheme was originally introduced to the GP context by \cite{williams2001using}. As discussed by \cite{quinonero2005unifying}, the Nystr\"om method by \cite{williams2001using} does not correspond to a well-formed probabilistic model. However, several methods modifying the inducing point approach are widely used. The Subset of Regressors (SoR) \citep{smola2001sparse} method uses the Nystr\"om approximation scheme and a finite linear-in-the-parameters model for approximating the whole (training and test) covariance function, whereas the sparse Nystr\"om method \citep{williams2001using} only replaces the training data covariance matrix. The SoR method is based on a degenerate prior which produces unreasonable predictive uncertainties, which is a general problem of linear models (for more details see \cite{rasmussen2006gaussian}).

@@

\vspace{2mm}
$\bullet$ In conventional sparse GP approximations, although the rank of the GP is reduced considerably to the number of inducing points, this still needs to do the autodiff and covariance matrix inversion.

\vspace{2mm}
$\bullet$ The Splines models are also a sort of basis functions expansion model, then the computational demands are similar to that in this approach. However in Splines models the lengthscale hyperparameter tend to be fixed and then the fit is covered by the magnitude parameter. In that sense, Splines models tend to loose the useful interpretation of the lengthscale parameter.

@@

\vspace{2mm}
$\bullet$ We investigate the relations going on among these factors, the number of basis functions, the box size, and the lengthscale of the functions.

\vspace{2mm}
$\bullet$ We make recommendations for the values of these factors based on the recognized relations among them. We provide useful graphs of these relations that will help the users to improve performance and save time of computation.

@@

Assuming that the spectral density function $S(\cdot)$ is regular enough, then it can be represented as a polynomial expansion:
%
\begin{equation}
S(\mathbf{w})=a_0+a_1\mathbf{w}^2+a_2(\mathbf{w}^2)^2+a_1(\mathbf{w}^2)^3+\cdots
\end{equation}
